\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2013-04-19 doc/sw-system.tex
agent-based modeling: latex and python now consistent

@@

\item How quickly can lenses be detected, and non-lenses be rejected?
How many classifiers are needed per target, and how quickly can they
mobilise?

@@

\section{Probabilistic Classification Model}
\label{appendix:probmodel}

>>>>

\sw is a web-based service
 that enables the discovery of strong
gravitational lenses in wide-field imaging surveys by arbitrarily large
numbers of people. Carefully produced color composite images are
displayed to volunteers via a flexible interface, which records
their estimates of the positions of candidate lensed features. Simulated
lenses, and expert-classified non-lenses, are inserted into the stream
at random intervals; this training set is used to give the volunteers
feedback on their performance as well as estimate a dynamically-updated
probability for any given image containing a lens. High probability
systems are filtered into the Zooniverse \Talk discussion system; this
smaller sample is further classified and analysed by volunteers into a
final set. We analyze the classification of the training set itself, and
find that 1) $<$volunteers learn$>$... 2) $<$the filter works$>$...

@@

\item How quickly can lenses be detected, and non-lenses be rejected?
How many volunteers are needed per target, and how quickly can they
mobilise?

@@

\section{Probabilistic Classification Data Model}
\label{appendix:probmodel}

Our aim is to enable the construction of a sample of good lens candidates.
Since we aspire to making logical  decisions, we define a  ``good candidate''
as one which has a high posterior probability of being a lens, given the data:
$\pr(\LENS|\data)$. Our problem is to approximate this probability. The data
in our case are the pixel values of a colour image. However, we can greatly
compress these complex, noisy sets of data by asking each volunteer what they
think about them. A complete  classification in \sw consists of a set of
Marker positions, or none at all. The null set encodes the statement from
the volunteer that the image in question is $\saidNOT$ a lens, while the
placement of any  Markers indicates that the volunteer considers this image to
contain a $\saidLENS$.  We simplify the problem by only using the Marker
positions to assess whether the volunteer  correctly assigned the
classification $\saidLENS$ or $\saidNOT$ after viewing (blindly) a member of
the training set of subjects.

How should we model these compressed data? The circumstances of each
classification are quite complex: the volunteers learn more about the problem
as they go, but also inevitably make occasional mistakes (perhaps because a
lens is difficult to see, or they became distracted by the television). To
cope with this uncertainty, we assign a software agent to partner each
volunteer. The agent's task is to interpret their volunteer's classification
data as best it can, using a model that makes a number of necessary
approximations. These interpretations will then include uncertainty arising as
a result of the volunteer's efforts and also the agent's approximations, but
they will have two important redeeming features. First, the interpretations
will be quantitative (where before they were qualititative),  and thus will be
useful in decision-making. Second, the agent will be able to predict, using
its model, the probability of a test subject being a $\LENS$, given both its
volunteer's classification, and its volunteer's experience. We now
describe how each agent works.

Each agent assumes that the probability of a volunteer recognising any given
simulated lens as a lens is some number, $\pr(\saidLENS|\LENS,T)$, that
depends only on what the volunteer is currently looking at, and all the
previous training subjects they have seen (and not on what type of lens it is,
how faint it is, what time it is, \etc). Likewise, it also assumes that the
probability of a volunteer recognising any given dud image as a dud is some
other number, $\pr(\saidNOT|\NOT,T)$, that also depends only on what the volunteer is currently looking at, and all the
previous training subjects they have seen. These two probabilities define a 
2 by 2 ``confusion matrix,'' which the agent updates, every time a
volunteer classifies a training subject, using the following 
very simple estimate:
\be
  \pr(``X"|X,T) \approx \frac{N_{``X"}}{N_X}.
  \label{eq:app:fraction}
\ee
Here, $X$ stands for the true classification of the subject, \ie either
$\LENS$ or $\NOT$, while $``X''$ is the corresponding classification
made by the volunteer on viewing the subject. $N_X$ is the number of
lenses the volunteer has been shown, while $N_{``X"}$ is the number of 
times the volunteer got their classifications of this type of training subject
right. $T$ stands for all
$N_{\LENS} + N_{\NOT}$ training data that the agent has heard about to
date.

The full confusion matrix of the $k^{\rm th}$ volunteer's agent is therefore:
\be
  \mathcal{M}^k = 
  \begin{bmatrix}
    \pr(\saidLENS|\NOT,T_k) & \pr(\saidLENS|\LENS,T_k) \\
    \pr(\saidNOT |\NOT,T_k) & \pr(\saidNOT |\LENS,T_k)
  \end{bmatrix}.
\ee
Note that these probabilities are normalized, such that
$\pr(\saidNOT |\NOT) = 1 - \pr(\saidLENS|\NOT)$.

Now, when this volunteer views a test subject, 
it is this confusion matrix that will allow their agent to update the
probability of that test subject being a $\LENS$. Let us suppose that
this subject has never been seen before: the agent assigns a 
prior probability that it is (or contains) a lens is 
\be
  \pr(\LENS) = p_0
\ee
where we have to assign a value for $p_0$. In the CFHTLS, we might expect
something like 100 lenses in 430,000 images, so $p_0 = 2\times10^{-4}$
is a reasonable estimate. The volunteer then makes a classification $C_k$ 
($= \saidLENS$ or $\saidNOT$).
We can apply Bayes' Theorem to derive how the agent should
update this prior probability into a posterior one using this new information:
\begin{align}
  \label{eq:app:first}
  & \pr(\LENS|C_k,T_k) = \\
  & \frac{\pr(C_k|\LENS,T_k)\cdot\pr(\LENS)}
{\left[ \pr(C_k|\LENS,T_k)\cdot\pr(\LENS) + \pr(C_k|\NOT,T_k)\cdot\pr(\NOT) \right]},
  \notag
\end{align}
which can be evaluated numerically using the elements of the confusion
matrix.

As an example, suppose we have a volunteer who is always right about the true
nature of a training subject. 
Their agent's confusion matrix would be
\be
  \mathcal{M}^{\rm perfect} = 
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$,
this hypothetical volunteer would submit $C = \saidLENS$. 
Their agent would then calculate the posterior probability for the 
subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier. 
Meanwhile, a hypothetical volunteer who (for some reason) 
wilfully always submits the wrong 
classification would have an agent with the column-swapped confusion matrix
\be
  \mathcal{M}^{\rm obtuse} = 
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful
information, since given the above confusion matrix, their agent would
calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers are as helpful as {\it perfect} ones!

Indeed, the information content of each classification can be estimated by an
agent before the next classification is made, just from its confusion matrix.
We define the expected information per classification, or {\it expertise}, as
\be
  \langle I_k \rangle = \frac{1}{2} 
                        \sum_{i} \left( \mathcal{M}^k_{i}  
              + \mathcal{M}^k_{i}\cdot\log_2{\mathcal{M}^k_i} \right),
  \label{eq:app:info}
\ee
where $i$ runs over all 4 elements of the confusion matrix 
$\mathcal{M}^k_i$, and $\langle I_k \rangle$ is measured in ``bits.'' 
The second term in the sum is recognisable as a
negative entropy, while the first term and the factor of 1/2
account for the way the confusion matrix encodes the results for the
classification of two types of subject, and that the probabilities
associated with each type are normalised. \Eref{eq:app:info} gives the
required result, that 
both the hypothetical {\it perfect} and
{\it obtuse} classifiers contribute 
1 bit of information each, per classification.

In fact, the only classifier that contributes the minimum information per
classification, zero bits, is the {\it random} classifier, whose agent's
confusion matrix elements are all 0.5.   
We conservatively initialise all the agents' confusion matrices to be that of
the random classifier. This makes no allowance for volunteers that actually do
have previous experience of what gravitational lenses look like, but should
help prevent large numbers of false positives being assigned high probability.
Plotting  $\langle I_k \rangle$ as a function of time will, to some extent,
illustrate the
learning process undergone by the $k^{\rm th}$ volunteer-agent partnership.

Suppose the $k+1^{\rm th}$ volunteer now submits a classification, on the same
subject just classified by the $k^{\rm th}$ volunteer. We can generalise
\Eref{eq:app:first} by replacing the prior probability with the current
posterior probability:
\begin{align}
  \label{eq:app:update}
  \pr(\LENS & |C_{k+1},T_{k+1},\data) = \\
  & \frac{1}{Z} \pr(C_{k+1}|\LENS,T_{k+1}) \cdot \pr(\LENS|\data) \\ \notag
{\rm where}\;\; Z = & \pr(C_{k+1}|\LENS,T_{k+1})\cdot\pr(\LENS|\data) \\ \notag
      & + \pr(C_{k+1}|\NOT,T_{k+1})\cdot\pr(\NOT|\data), \notag
\end{align}
and $\data = \{C_k,T_k\}$ is the set of all previous
classifications, and the set of training subjects seen by each of those
volunteers.
$\pr(\LENS|\data)$ is the fundamental property of each test subject that
we are trying to infer. We track $\pr(\LENS|\data)$ as a function of time,
and by comparing it to a lower or upper thresholds, make decisions about
whether to retire the subject from the classification interface or
promote it in \Talk, respectively.
